---
title: "Breast Cancer Detection with Heterogeneous Ensemble Classifier"
author: "Your Name"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
---

# Breast Cancer Detection with Heterogeneous Ensemble Classifier

## Introduction

This project aims to develop a heterogeneous ensemble classifier for breast cancer detection using the BreastCancer dataset. The ensemble will combine predictions from multiple classifiers trained using different algorithms.

## Dataset

We'll start by loading the BreastCancer dataset provided by the `mlbench` package and inspecting its structure.

```{r load_dataset, message=FALSE}
library(mlbench)
data(BreastCancer)
str(BreastCancer)
```

## Data Preprocessing

Before training our classifiers, it's essential to preprocess the data by handling missing values and removing unnecessary columns.

```{r data_preprocessing}
# Handle missing values
BreastCancer <- na.omit(BreastCancer)

# Remove the unique identifier
BreastCancer$Id <- NULL
```

## Exploratory Data Analysis (EDA)

Let's explore the distribution of classes and visualize relationships between features.

### Class Distribution

```{r class_distribution}
library(ggplot2)
ggplot(BreastCancer, aes(x = Class, fill = Class)) +
  geom_bar() +
  labs(title = "Class Distribution", x = "Class", y = "Frequency")
```

### Scatter Plot

```{r scatter_plot}
ggplot(BreastCancer, aes(x = Cl.thickness, y = Cell.size, color = Class)) +
  geom_point() +
  labs(title = "Scatter plot of Cl.thickness vs Cell.size", x = "Cl.thickness", y = "Cell.size", color = "Class")
```

### Box Plot

```{r box_plot}
ggplot(BreastCancer, aes(x = Class, y = Cell.shape, fill = Class)) +
  geom_boxplot() +
  labs(title = "Box plot of Cell.shape by Class", x = "Class", y = "Cell.shape")
```

## Model Training

We will train classifiers using various algorithms:

```{r train_classifiers}
library(rpart)
library(klaR)
library(nnet)
library(randomForest)

### Train Decision Tree
classifier_rpart <- rpart(Class ~ ., data = BreastCancer)

### Train Naive Bayes
classifier_nb <- NaiveBayes(Class ~ ., data = BreastCancer)

### Train Neural Network
classifier_nnet <- nnet(Class ~ ., data = BreastCancer, size = 1)

### Train Linear Discriminant Analysis
classifier_lda <- lda(Class ~ ., data = BreastCancer)

### Train Random Forest
classifier_rf <- randomForest(Class ~ ., data = BreastCancer)
```

## Ensemble Classification

In this section, we'll combine predictions from the trained classifiers using the Majority Rule ensemble approach.

```{r ensemble_classification}
# Make predictions from each classifier and suppress warnings
pred_rpart <- suppressWarnings(try(predict(classifier_rpart, newdata = BreastCancer, type = "class"), silent = TRUE))
pred_nb <- suppressWarnings(try(predict(classifier_nb, newdata = BreastCancer), silent = TRUE))
pred_nnet <- suppressWarnings(try(predict(classifier_nnet, newdata = BreastCancer, type = "class"), silent = TRUE))
pred_lda <- suppressWarnings(try(predict(classifier_lda, newdata = BreastCancer)$class, silent = TRUE))
pred_rf <- suppressWarnings(try(predict(classifier_rf, newdata = BreastCancer), silent = TRUE))

# Check for errors in predictions
if (class(pred_rpart) == "try-error" || class(pred_nb) == "try-error" || 
    class(pred_nnet) == "try-error" || class(pred_lda) == "try-error" || 
    class(pred_rf) == "try-error") {
  cat("Error: Unable to make predictions from one or more classifiers.\n")
} else {
  # Combine predictions using Majority Rule
  ensemble_predictions <- ifelse(
      rowSums(cbind(pred_rpart == "malignant", pred_nb == "malignant", 
                    pred_nnet == "malignant", pred_lda == "malignant", 
                    pred_rf == "malignant")) >= 3, 
      "malignant", "benign")
  
  # Calculate accuracy
  accuracy <- mean(ensemble_predictions == BreastCancer$Class)
  
  # Display ensemble predictions for the first 10 rows
  head(ensemble_predictions, 10)
  
  # Visualize ensemble predictions
  library(ggplot2)
  ggplot(data.frame(predictions = ensemble_predictions), aes(x = predictions)) +
    geom_bar() +
    labs(title = "Ensemble Predictions", x = "Predicted Class", y = "Count")
}

```

## Performance Evaluation

Let's evaluate the performance of our ensemble classifier using a confusion matrix.

```{r confusion_matrix}
library(caret)

# Convert predicted classes to factors with the same levels as actual classes
ensemble_predictions_factor <- factor(ensemble_predictions, levels = levels(BreastCancer$Class))

# Convert actual classes to factors with the same levels as predicted classes
actual_classes_factor <- factor(BreastCancer$Class, levels = levels(ensemble_predictions_factor))

# Calculate confusion matrix
conf_matrix <- confusionMatrix(ensemble_predictions_factor, actual_classes_factor)
conf_matrix
```

## Conclusion

In this project, we developed a heterogeneous ensemble classifier for breast cancer detection. By combining predictions from multiple classifiers using the Majority Rule ensemble approach, we achieved an accuracy of `r accuracy`%. The confusion matrix provides insights into the performance of our classifier.
```
